### 2021-12-15 Draft Final Project1.Rmd
Add files via upload

---
title: "Final Project Data 205"
author: "Tiffany King"
date: "11/18/2021"
output: html_document
---

# Loading Libraries
```{r}
library('rvest')
library(tidyverse)
library(psych)
library(readr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(plotly)
library(plyr)
library(highcharter)
library(RColorBrewer)
```
# Setting working directory
```{r}
setwd("/Users/tiffanyking/Desktop/DATA 205/DATA205/DATA205")
```

```{r}
WKridership<- read_csv("2019 Ridership weekday average.csv")
crimes<-read_csv("crime.csv")
``` 
# Examine the data
```{r}
str(WKridership)
head(WKridership)
```
## create dataframe from CSV
```{r}
year<- c("2010","2011","2012","2013","2014","2015","2016","2017","2018","2019","2020","2021")
averider<- c(705563,737313,721225,706591,693705,675077,621079,613665,605910,626269,176602,125557)
dailyentriesride<-data.frame(year,averider)
print(dailyentriesride)
```
# Barplot Average Daily Entries 
```{r}
options(scipen = 999)
ggplot(data=dailyentriesride, aes(x=year, y=averider)) +
  geom_bar(stat="identity", fill="red")+
  geom_text(aes(label=averider), vjust=1.6, color="white", size=3.5)+
   xlab("Year") +
  ylab("Passengers") +
  ggtitle("Average Daily Entries By Train") +
  ylim(0,800000)
  theme_minimal()
```

## Getting the WMATA Data
```{r}
library(tidyverse)
library(pdftools)
library(httr)
```

```{r}
get_wmata_data <- function(pdflink) {
  wmata_data <- pdf_text(pdflink) %>%
    str_split("\n")

  df = ""

  for (i in 1:length(wmata_data)) {

    for (line in wmata_data[[i]][9:(length(wmata_data[[i]])-2)]) {
      if (line == "" | is.na(line))
      {
        next
      } else {
        line <- str_squish(line)

        if (substr(line,21,27) %in% c("SUMMONS"))
        {
          stub <- str_split_fixed(line, " - ", 2)
          state <- substr(stub[1], nchar(stub[1])-1, nchar(stub[1]))
          stub_1 <- substr(stub[1],1, nchar(stub[1])-3)
          stub_2 <- stub[2]
          line <- paste(stub_1, stub_2, "\n")
          line <- paste(substr(line, 1, 36), state, " ", substr(line, 37, nchar(line)))
          prev_line <- line
          #print(line)
        }

        if (substr(line,21,27) %in% c("ARREST ", "REPORT "))
        {
          line <- paste(substr(line,1,26), "        ", substr(line,27,nchar(line)))
          stub <- str_split_fixed(line, " - ", 2)
          state <- substr(stub[1], nchar(stub[1])-1, nchar(stub[1]))
          stub_1 <- substr(stub[1],1, nchar(stub[1])-3)
          stub_2 <- stub[2]
          line <- paste(stub_1, stub_2, "\n")
          line <- paste(substr(line, 1, 36), state, " ", substr(line, 37, nchar(line)))
          prev_line <- line
          #print(line)
        }

        if (substr(line,3,3) !="/")
        {
          offense_stub <- str_split_fixed(line, " - ", 2)

          if (offense_stub[2]  != "")
          {
            line <- paste(stub_1, substr(line,6,nchar(line)))
            line <- paste(substr(line, 1, 36), state, " ", substr(line, 37, nchar(line)), "\n")
          }else{
            df <- substr(df, 1, nchar(df)-nchar(prev_line))
            line <- paste(substr(prev_line,1,nchar(prev_line)-2), offense_stub[1], "\n")
          }
        }

        df <- paste(df, line)
        print(line)
      }
    }
  }
  return(read_fwf(df, 
                  col_types=cols('c','c','c','c','c'),
                  fwf_cols(date=12, time=9, type=17, state=4, desc=NA)))
}
#fwf_cols(date=12, time=9, type=17, state=4, desc=NA),
# Now define what the links to the pdf files should look like and create a
# way to iterate through them to get multiple files at one go.

url <- "https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-"
time_periods <- c("October-2021","September-2021", "August-2021", "July-2021", "June-2021","May-2021","April-2021","March-2021","February-2021","January-2021","December-2020","November-2020","October-2020","September-2020","August-2020","May-2020","April-2020","March-2020","February-2020","January-2020","December-2019","November-2019","October-2019","June-2019","November-2018","March-2018","February-2018","January-2018","October-2017","September-2017")
```

```{r}
url_exception_june_2020<-"https://www.wmata.com/about/transit-police/upload/June-2020-Blotter-Report.pdf"
df_exception_June_2020 <- get_wmata_data(url_exception_june_2020)

url_exception_july_2020<-"https://www.wmata.com/about/transit-police/upload/July-2020-Blotter.pdf"
df_exception_July_2020 <- get_wmata_data(url_exception_july_2020)

url_exception_september_2019<-"https://www.wmata.com/about/transit-police/upload/September-2019-Monthly-Blotter.pdf"
df_exception_September_2019 <- get_wmata_data(url_exception_september_2019)

url_exception_august_2019<-"https://www.wmata.com/about/transit-police/upload/August-2019-Monthly-Blotter.pdf"
df_exception_August_2019 <- get_wmata_data(url_exception_august_2019)

url_exception_july_2019<-"https://www.wmata.com/about/transit-police/upload/July-2019-Blotter.pdf"
df_exception_July_2019 <- get_wmata_data(url_exception_july_2019)

url_exception_may_2019<-"https://www.wmata.com/about/transit-police/upload/May-2019-Blotter-Report.pdf"
df_exception_May_2019 <- get_wmata_data(url_exception_may_2019)

url_exception_april_2019<-"https://www.wmata.com/about/transit-police/upload/Blotter-Report-April-2019.pdf"
df_exception_April_2019 <- get_wmata_data(url_exception_april_2019)

url_exception_march_2019<-"https://www.wmata.com/about/transit-police/upload/2019-March-Blotter-Report.pdf"
df_exception_March_2019 <- get_wmata_data(url_exception_march_2019)

url_exception_february_2019<-"https://www.wmata.com/about/transit-police/upload/February-2019-Monthly-Blotter-Report.pdf"
df_exception_February_2019 <- get_wmata_data(url_exception_february_2019)

url_exception_january_2019<-"https://www.wmata.com/about/transit-police/upload/January-2019-Monthly-Blotter-Report.pdf"
df_exception_January_2019 <- get_wmata_data(url_exception_january_2019)

url_exception_december_2018<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-2018-December.pdf"
df_exception_December_2018 <- get_wmata_data(url_exception_december_2018)

url_exception_september_2018<-"https://www.wmata.com/about/transit-police/upload/September-2018-Blotter.pdf"
df_exception_September_2018 <- get_wmata_data(url_exception_september_2018)

url_exception_august_2018<-"https://www.wmata.com/about/transit-police/upload/August-2018-Monthly-Blotter.pdf"
df_exception_August_2018 <- get_wmata_data(url_exception_august_2018)

url_exception_july_2018<-"https://www.wmata.com/about/transit-police/upload/July-2018-Blotter.pdf"
df_exception_July_2018 <- get_wmata_data(url_exception_july_2018)

url_exception_june_2018<-"https://www.wmata.com/about/transit-police/upload/June-2016-Blotter.pdf"
df_exception_June_2018 <- get_wmata_data(url_exception_june_2018)

url_exception_may_2018<-"https://www.wmata.com/about/transit-police/upload/May-Monthly-blotter-2018.pdf"
df_exception_May_2018 <- get_wmata_data(url_exception_may_2018)

url_exception_april_2018<-"https://www.wmata.com/about/transit-police/upload/April-Monthly-blotter.pdf"
df_exception_April_2018 <- get_wmata_data(url_exception_april_2018)

url_exception_december_2017<-"https://www.wmata.com/about/transit-police/upload/December-Monthly-Blotter.pdf"
df_exception_December_2017 <- get_wmata_data(url_exception_december_2017)

url_exception_november_2017<-"https://www.wmata.com/about/transit-police/upload/November-Monthly-Blotter.pdf"
df_exception_November_2017 <- get_wmata_data(url_exception_november_2017)

url_exception_august_2017<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-Report-August-2017.pdf"
df_exception_August_2017 <- get_wmata_data(url_exception_august_2017)

url_exception_july_2017<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter-Report-July-2017.pdf"
df_exception_July_2017 <- get_wmata_data(url_exception_july_2017)

url_exception_june_2017<-"https://www.wmata.com/about/transit-police/upload/June-2017-blotter.pdf"
df_exception_June_2017 <- get_wmata_data(url_exception_june_2017)

url_exception_april_2017<-"https://www.wmata.com/about/transit-police/upload/April-2017-monthly-blotter.pdf"
df_exception_April_2017 <-get_wmata_data(url_exception_april_2017)

url_exception_march_2017<-"https://www.wmata.com/about/transit-police/upload/March-2017-monthly-blotter.pdf"
df_exception_March_2017 <-get_wmata_data(url_exception_march_2017)

url_exception_february_2017<-"https://www.wmata.com/about/transit-police/upload/January-2017-monthly-blotter.pdf"
df_exception_February_2017 <-get_wmata_data(url_exception_february_2017)

url_exception_january_2017<-"https://www.wmata.com/about/transit-police/upload/Monthly-Blotter_February-2017.pdf"
df_exception_January_2017 <-get_wmata_data(url_exception_january_2017)

df_prev = ""

for (time_period in time_periods) {
  pdflink <- paste0(url, time_period, ".pdf")
  df_curr <- get_wmata_data(pdflink)

  if (df_prev == "")
  {
    df_prev <- df_curr
    df_wmata <- df_prev
  } else {
    df_wmata <- bind_rows(df_wmata,df_curr,df_exception_June_2020,df_exception_July_2020,df_exception_September_2019,df_exception_August_2019,df_exception_March_2019,df_exception_February_2019,df_exception_January_2019,df_exception_December_2018,df_exception_September_2018,df_exception_August_2018,df_exception_July_2018, df_exception_June_2018,df_exception_May_2018,df_exception_April_2018,df_exception_December_2017,df_exception_November_2017,df_exception_August_2017,df_exception_July_2017,df_exception_June_2017,df_exception_April_2017,df_exception_March_2017,df_exception_February_2017,df_exception_January_2017)

  }
}

# Check to make sure the data starts with start month and ends with end month.
df_wmata
tail(df_wmata)

```

## View df_wmata Data 
```{r}
view(df_wmata)
```

## Summary
```{r}
summary(df_wmata)
```

## Get rid of NAs
```{r}
WMATA <-na.omit(df_wmata)
```

## Rename variables (State)
```{r}
WMATA$state[grepl('PG',WMATA$state)]<- 'MD'
WMATA$state[grepl('MC',WMATA$state)]<- 'MD'
WMATA$state[grepl('RL',WMATA$state)]<- 'VA'
WMATA$state[grepl('EX',WMATA$state)]<- 'VA'
```

## Separate year from date
```{r}
WMATA$date <- as.Date(WMATA$date, format="%m/%d/%Y")
head(WMATA)      
```

```{r}
WMATA$year<-as.numeric(format(WMATA$date,"%Y"))
WMATA$month<-as.numeric(format(WMATA$date,"%m"))
head(WMATA)
```

## Groupby Time Series( Define Time (Morning, Afternoon ,Evening, Night)

```{r}
#Create a time series plot
WMATA$hour <- as.numeric(gsub(":.*$", "", WMATA$time))
WMATA$timeofday <- ifelse(WMATA$hour >= 5 & WMATA$hour<=11, "Morning",
                            ifelse(WMATA$hour > 11 & WMATA$hour<=16, "Afternoon",
                            ifelse(WMATA$hour >16 & WMATA$hour <=19, "Evening", "Night"))) 
WMATA
```

## Plot 1
```{r}
p1<-WMATA  %>%
  filter(state %in% c("DC","MD","VA")) %>%
  ggplot(aes(x=year, fill= state))+
  geom_bar()+
  labs(title="Total of Crimes by State/District",x="State/District", y="Total")+
  scale_fill_brewer(palette = "BuPu") + theme_minimal()
p1 + theme_minimal()
```

## Plot Types of Crime
```{r}
p2<-WMATA  %>%
  filter(type %in% c("ARREST","REPORT","SUMMONS/CITATION")) %>%
  ggplot(aes(x=type, fill= type))+
  geom_bar()+
  labs(title="Classification of Type of Crime/ Reporting",x="Type", y="Total")+
  scale_fill_brewer(palette = "") + theme_minimal()
p2 + theme_minimal()
```

## Preparing the data (time)
```{r}
d3<- WMATA%>%
  group_by(year, timeofday) %>%
  dplyr::summarize(Freq=n())
```

##Fix Na's 
```{r}
d4 <-na.omit(d3)
```

## Facet
```{r}
 ggplot(data = d4,
        mapping = aes(x = year, y = Freq, fill=timeofday)) +
     geom_bar(stat="identity") +
     facet_wrap(vars(timeofday)) +
      scale_fill_brewer(palette = "Spectral") +
     theme_minimal()

```

##Prepare the Data
```{r}
DMV<-WMATA %>%
  select(type,year)%>%
  group_by(year, type)%>%
  filter(year >=2017  & type %in% c("ARREST","REPORT","SUMMONS/CITATION"))%>%
  dplyr::summarize(Freq=n())
```

```{r}
head(DMV)
```
```{r}
ggplot(data = DMV,
      mapping = aes(x = year, y = Freq)) +
     geom_line(stat="identity") +
     facet_wrap(vars(type)) +
     scale_fill_brewer(palette = "Spectral") +
     theme_minimal()
```

## Months & Count
```{r}
DF6<- WMATA%>%
  group_by(month) %>%
  dplyr::summarize(count=n())
DF6
```

```{r}
df7<-na.omit(DF6)
```

## Time Series 
```{r}
P6<- ggplot(df7, aes(x=month, fill=month)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2,
                 position=position_dodge(.9))
```

## Getting the WMATA Geocodes
```{r}
library(dplyr)
library(httr)
#library(jsonlite)
library(stringr)
library(tidygeocoder)

# Extract the street address from desc using Regex.  Extract crime description also.
df_wmata2 <- WMATA %>% mutate(street_addr=str_remove(desc, "[[:upper:]]{2,}.+")) %>%
  mutate(crime_desc=str_extract(desc, "[[:upper:]]{2,}.+"))
df_wmata2 <- df_wmata2 %>% mutate(full_addr=paste(street_addr,state))

#Filter on just Maryland addresses
df_md <- df_wmata2 %>% filter(state=="MD")
```

```{r}
#Apply the geocode function to get lat/lon values
coordinates <- df_md %>% geocode(full_addr, method="arcgis")

coordinates$census_tract <- 0
row_count <- nrow(coordinates)

# Get FIPS code from FCC API based on lat/lon
fcc_url = "https://geo.fcc.gov/api/census/block/find?"
headers = "'Content-Type':'application/json'"
unique_record = 0

for (row in 1:row_count) {
  if (!is.na(coordinates$lat[row]) & coordinates$time[row] != unique_record) {

    full_url <- paste0(fcc_url, "latitude=", coordinates$lat[row],
                       "&longitude=", coordinates$long[row], "&censusYear=2010&format=json")
    fcc <- GET(full_url)
    fcc_parsed <- content(fcc, as="parsed")
    try(coordinates$census_tract[row] <- fcc_parsed$Block$FIPS, TRUE)

    print(full_url)
    print(coordinates$census_tract[row])
    unique_record = coordinates$time[row]
  }
  Sys.sleep(1)
}

census_info <- GET("https://api.census.gov/data/2019/acs/acs5?get=NAME,B01003_001E,B02001_002E,B02001_003E&for=block%20group:*&in=state:24&in=county:031&in=tract:*")
df_census <- fromJSON(content(census_info, as="text"))

df_census
```

```{r}
P5<- ggplot(DMV, aes(y =year, fill = factor(type))) +
  geom_boxplot() +
  facet_wrap(~type, scale = "free") +
  scale_y_continuous(name = "Average Life Expectancy") +
  scale_x_discrete(labels = abbreviate, name = "Continent")

P5
```

```{r}
ggplot(DMV, aes(x=type)) + geom_histogram()
```

```{r}
p <- ggplot(DMV, aes(year, type)) + 
  geom_abline() +
  geom_jitter(width = 0.1, height = 0.1) 
p + facet_wrap()
```

## Facet_wrap on the types of reports throughout the years
```{r}
ggplot(data = WMATA, aes(year, n)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "steelblue") + 
  labs(title = "New Marvel characters by alignment",
       subtitle = "(limited to characters with more than 100 appearances)",
       y = "Count of new characters", x = "") + 
  facet_grid(~type==c("DC","MD","VA"))
```

## Two links that will not load & has an error
```{r}

url_exception_october_2018<-"https://www.wmata.com/about/transit-police/upload/October-2018-Monthly-Crime-Blotter.pdf"
df_exception_October_2018 <- get_wmata_data(url_exception_october_2018)

url_exception_may_2017<-"https://www.wmata.com/about/transit-police/upload/May-2017-Blotter.pdf"
df_exception_May_2017 <- get_wmata_data(url_exception_may_2017)

```

#Census Data 
```{r}
library(httr)
library(jsonlite)
library(tidyverse)

# Import json data from Census API
data <- GET("https://api.census.gov/data/2019/acs/acs1/?get=NAME,B01001_001E,B01002_001E,C22001_001E,B25064_001E&for=county:*&in=state:24")
data <- content(data, as="text")

# Load into a data frame
df <- fromJSON(data)
class(df)
head(df)

# Change to tibble
df <- as_tibble(df)
head(df)

# Make first row column header names
names(df) <- c(df[1,])
head(df)

# Keep all but first row (which had the column header names)
df <- df[-1,]
head(df)

# Rename columns as needed
names(df) <- c('NAME',
               'B01001_001E' = 'Population',
               'B01002_001E' = 'Median_Age',
               'C22001_001E' = 'Household_Poverty',
               'B25064_001E' = 'Median_Gross_Rent',
               'state',
               'county')
head(df)

```

```{r}
library(dplyr)
library(httr)
#library(jsonlite)
library(stringr)
library(tidygeocoder)

# Extract the street address from desc using Regex.  Extract crime description also.
df_wmata2 <- WMATA %>% mutate(street_addr=str_remove(desc, "[[:upper:]]{2,}.+")) %>%
  mutate(crime_desc=str_extract(desc, "[[:upper:]]{2,}.+"))
df_wmata2 <- df_wmata2 %>% mutate(full_addr=paste(street_addr,state))

#Filter on just Maryland addresses
df_md <- df_wmata2 %>% filter(state=="MD")

#Apply the geocode function to get lat/lon values
coordinates <- df_md %>% geocode(full_addr)

coordinates$census_tract <- 0
row_count <- nrow(coordinates)
```

## Create on plot on Income & Population(Linear Regression/ Scatterplot with Population)
```{r}

```

## Remove Crimes Data 
```{r}
Moco_crimes<-na.omit(crimes)
```

## Plot on Classifcation of Crime
```{r}
c1<-crimes  %>%
  filter(`Crime Name1` %in% c("Crime Against Person","Crime Against Property","Crime Against Society")) %>%
  ggplot(aes(x=`Crime Name1`, fill=`Crime Name1`))+
  geom_bar()+
  labs(title="Classification of Crime",x="", y="Total")+
  scale_fill_brewer(palette = "Accent") + theme_minimal()
c1 + theme_minimal()
```

## f
```{r}

```